\section{Research Questions and Study Design}

The study is organized around four research questions that connect the modeling choices to measurable outcomes.

\textbf{RQ1: How can WLAN mobility traces be converted into a cooperative caching topology?}
We model APs as graph nodes and user handoffs as weighted edges. The resulting graph captures observed mobility-induced interaction strength rather than assuming that every AP has the same cooperation neighborhood.

\textbf{RQ2: Does topology-aware placement improve over local and global popularity baselines?}
We compare topology-aware placement against random placement, local popularity, global popularity, and diversified popularity. This separates the value of graph-aware optimization from the value of simple catalog diversity.

\textbf{RQ3: Does learning-based refinement add value beyond topology-aware greedy placement?}
The DQN refiner starts from the currently deployed placement and searches a local neighborhood of node-content changes. This tests whether learning can improve a strong structured cache state without requiring a neural policy to learn the full placement problem from scratch.

\textbf{RQ4: Is transition-aware online refinement sufficient under sequential topology perturbation?}
We evaluate policies across multiple perturbation rates as an online sequence. This tests whether online DQN should keep refining the deployed cache as AP availability and cooperative edges change, and whether a lightweight validation gate actually changes the deployed strategy once no-op, diversified, greedy, and DQN candidates are compared with relocation and optimization cost included.

\subsection{Research Hypotheses}

The research questions lead to three hypotheses. First, topology-aware placement should outperform local and global popularity because it can exploit cooperative paths while controlling duplication. Second, transition-aware DQN refinement should improve a deployed placement when the current cache state is strong but not locally optimal. Third, a validation gate may improve operating efficiency if cheap candidates or no-op become preferable under some perturbation regimes; if it never changes the decision, the evidence should instead support a simpler conclusion that online DQN refinement is sufficient for the tested regimes.

These hypotheses are intentionally modest. The study does not assume that DQN is useful merely because it is neural, that diversity is always best, or that topology awareness alone solves every perturbation condition. Instead, the experimental design asks whether each behavior is useful after relocation is charged against the previous online cache state.

\subsection{Gap Closure Strategy}

Prior conceptual caching studies sometimes describe a graph, a greedy algorithm, and a learning component without specifying the operational details needed to reproduce the claims. This study closes those gaps by fixing the graph-construction rules, defining a stochastic objective, specifying the DQN state/action/reward design, using strong baselines, and reporting measured metrics under repeated perturbation samples.

\begin{table}[t]
\centering
\caption{Major methodology gaps addressed in this study.}
\label{tab:gap_closure}
\begin{tabular}{p{0.29\textwidth}p{0.61\textwidth}}
\toprule
Gap & Closure \\
\midrule
Unspecified cooperation graph & APs are represented as graph nodes; handoff counts define weighted edges and latency weights. \\
Ambiguous dataset handling & Official Dartmouth/IEEE DataPort access is documented, and trace-compatible mobility data is generated deterministically when raw traces are unavailable. \\
Unrealistic demand homogeneity & Demand uses Zipf popularity with AP-level spatial locality. \\
Single weak baseline comparison & Evaluation includes random, local, global, diversified, greedy, fixed DQN-refined, online DQN-refined, no-op, and validation-gated policies. \\
Conceptual learning component & DQN uses explicit state, action, reward, replay, target network, exploration, and accepted-action inference. \\
No robustness evaluation & Policies are evaluated over repeated node and edge perturbation samples. \\
Overclaiming one method & The validation gate is reported explicitly and interpreted as a safeguard, not as an independent source of improvement when it selects online DQN in all tested windows. \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Interpretation of Adaptation}

Adaptation in this study is not defined as arbitrary cache rewriting. It has a precise meaning. At the placement level, online DQN refinement proposes changes to the currently deployed placement and accepts only objective-improving actions. At the deployment level, the validation gate selects the candidate with the best cost-aware validation score under the current perturbation rate, including the option to keep the current placement. Therefore, adaptation occurs through bounded local refinement and explicit validation, not through uncontrolled replacement of all cached content.

Graph perturbation also has a concrete meaning. If a node is removed, cached objects stored only at that node are no longer available in the active component. If an edge is removed, shortest-path retrieval latency may increase and cooperative hits may become origin misses. If strongly connected APs remain active, redundant copies among them still contribute to redundancy cost. These changes explain why the same placement matrix can produce different access cost, hit decomposition, redundancy, and objective value under different graph samples.
