\section{Related Work}

\subsection{Caching Foundations}

Cache replacement and placement strategies have been studied for decades. Classical policies such as LRU, LFU, FIFO, and popularity-based replacement remain important because they are simple and practical, but they are usually designed for a single cache or for systems where cooperation is not modeled explicitly \cite{Podlipnig2003}. Zipf-like popularity has been observed in web caching and content distribution, and it remains a common basis for synthetic content demand models \cite{breslau1999zipf}. These foundations motivate the popularity baselines in this study, but they do not directly solve cooperative placement over a mobility-derived graph.

Survey work on cellular and wireless caching emphasizes that caching must be evaluated with respect to placement, delivery, latency, backhaul load, power, and system architecture \cite{Li2018CachingSurvey,Paschos2018RoleCaching}. This broader view is important because cache hit ratio alone can be misleading. A policy with a high hit ratio may still be undesirable if it creates excessive duplication, requires frequent updates, or places objects at nodes that are not reachable under perturbation. Our objective therefore includes access cost, replication cost, redundancy cost, and relocation cost.

\subsection{Cooperative Wireless and Edge Caching}

Cooperative caching studies show that coordinated placement across helpers can improve wireless content delivery. FemtoCaching demonstrates that helper caches can reduce download delay and backhaul usage when content placement is optimized jointly \cite{Shanmugam2013,Golrezaei2014JSAC}. Coded caching shows that coordinated placement and coded delivery can achieve theoretical efficiency gains in broadcast-style settings \cite{MaddahAli2014}. Small-cell caching formulations also show that caching and routing decisions are coupled and can be computationally difficult \cite{Poularakis2014Approx}.

These works establish the value of cooperation, but many assume known helper relationships, static topology, or a delivery model that differs from AP-level mobility. In contrast, this study derives the cooperation graph from handoff behavior and evaluates placements under node and edge perturbations. The difference is significant: in a mobility-derived graph, cooperation strength is not merely a physical-neighborhood assumption, and redundancy should be penalized more strongly among APs that exchange many users.

\subsection{Information-Centric and Popularity-Aware Edge Caching}

Information-centric networking and named-data networking place content names and in-network storage at the center of network design \cite{Zhang2014NDN}. Popularity-aware and closeness-based caching in NDN edge networks further shows that network position and demand popularity should be considered jointly \cite{Amadeo2022}. These ideas are closely related to topology-aware caching because they treat retrieval location and network distance as first-class design variables.

The present study follows the same principle but uses an AP handoff graph rather than an ICN forwarding topology. The graph is built from mobility transitions, and the retrieval cost is computed over shortest paths in the active perturbed graph. This makes the method appropriate for WLAN-like IoT edge environments where AP association behavior is observable even when application-layer content logs are unavailable.

\subsection{Mobility-Aware and Delay-Tolerant Caching}

Mobility-aware caching recognizes that user movement affects future demand and content availability. Mobile edge network surveys highlight the convergence of computing, caching, and communication in mobile systems \cite{Wang2017Mobility}. Delay-tolerant and opportunistic caching methods use contact patterns, mission context, or social context to decide where content should be placed \cite{Mardham2018,Cabaniss2014}. These methods demonstrate that mobility and contact structure can guide caching decisions.

The distinction in this study is that mobility is represented as an AP-level weighted graph and inserted directly into the objective. Instead of using mobility only to predict a future request or to define a social heuristic, the handoff graph affects access latency, redundancy cost, perturbation response, and policy selection. This creates a tighter connection between mobility structure and cache placement.

\subsection{Learning-Based Edge Caching}

Learning-based caching has been widely studied because demand, mobility, and network conditions are non-stationary. Early learning-based proactive caching uses online learning and optimization to adjust cache content under changing popularity \cite{Blasco2014}. Deep reinforcement learning has been applied to edge caching, vehicular caching, and self-driving or multi-access edge settings \cite{Chen2021TMC,Ndikumana2020,zhou2023distributed}. DQN and Double DQN provide the replay-buffer, target-network, and stabilized Q-learning mechanisms that motivate the learning component used here \cite{mnih2015dqn,vanhasselt2016double}.

Multi-agent reinforcement learning has also been used for cooperative edge caching, including bandwidth-efficient inter-agent communication and distributed decision making \cite{Min2020MARL,zhou2023distributed}. Such approaches are powerful, but they can be difficult to train when the action space is the full node-content placement matrix. Adaptive TACC uses a different design: topology-aware greedy produces a strong warm start, and DQN performs local refinement. This reduces the learning burden and makes the neural component easier to interpret.

\subsection{Recent Vehicular, UAV, Satellite, and Federated Caching}

Recent work extends cooperative caching to vehicular, UAV-assisted, and satellite-assisted edge systems. Bai et al. frame collaborative edge intelligence as a broader direction in which vehicles, edge nodes, and cloud resources coordinate to support low-latency autonomous services \cite{bai2025collaborative}. Xu et al. study cooperative edge content caching with popularity prediction in UAV-assisted vehicular networks \cite{xu2026cooperative}. Ma et al. propose delay-oriented caching for LEO satellite-assisted edge caching networks \cite{ma2025delay}. Cao et al. combine federated distillation and reinforcement learning for mobility-aware cooperative caching in vehicular edge computing \cite{cao2025}. Zhao et al. jointly consider collaborative caching and offloading for vehicular edge computing \cite{zhao2026collaborative}.

These recent studies confirm that caching is moving toward mobility-aware, collaborative, and learning-based edge systems. They also show that no single network setting fully captures the problem. Vehicular and UAV systems emphasize fast mobility, non-terrestrial systems emphasize long and variable transmission paths, and federated approaches emphasize privacy-preserving distributed learning. Our study focuses on WLAN-style AP handoff topology, where the central challenge is to convert observed AP transitions into a cooperative placement graph and evaluate cache decisions under graph perturbation.

Bandit and federated methods provide another approach to adaptation. Dhara et al. use contextual and federated multi-armed bandits for cooperative real-time content retrieval in integrated space-air-ground networks under disaster conditions \cite{dhara2026co}. Bandit methods are attractive when decisions decompose into smaller actions and rewards can be learned online. Our placement problem is a binary node-content matrix with capacity constraints, so we combine greedy search, DQN refinement, and validation-based selection rather than using a pure bandit formulation.

\subsection{Graph and Knowledge-Based Network Optimization}

Graph neural networks have strong appeal for network optimization because communication networks are relational objects rather than fixed-order vectors. GNNs have been used for SDN modeling and broader communication-network optimization \cite{Rusek2019,Jiang2022Survey}. Knowledge-graph methods have also been proposed for cooperative caching in MEC-enabled heterogeneous networks \cite{wang2024knowledge}. These methods suggest that graph representation can improve generalization across topologies and capture relational structure more naturally than hand-crafted features.

Adaptive TACC is graph-aware but not a full GNN method. We use explicit graph features, including degree, weighted degree, and closeness centrality, inside the DQN state, and the objective itself uses shortest-path latency and edge-weighted redundancy. This design keeps the method lightweight and reproducible while preserving a path toward future GNN extensions.

\subsection{Positioning}

Table~\ref{tab:related_comparison} summarizes the position of this study. The contribution is not that mobility, caching, or DRL are individually new. The contribution is their integration into a single topology-aware caching framework with perturbation-aware evaluation and adaptive selection among policy families. This combination is useful because recent literature increasingly recognizes mobility-aware caching, but many studies either assume a different network domain, train learning policies over large action spaces, or do not explicitly evaluate how perturbations change which placement family is best.

\begin{table}[t]
\centering
\caption{Comparison of related work themes and the contribution of this study.}
\label{tab:related_comparison}
\begin{tabular}{p{0.24\textwidth}p{0.30\textwidth}p{0.32\textwidth}}
\toprule
Research theme & Typical strength & Gap addressed here \\
\midrule
Classical cache replacement \cite{Podlipnig2003} & Simple online operation & Does not explicitly coordinate neighboring edge caches. \\
Femto/helper caching \cite{Shanmugam2013,Golrezaei2014JSAC} & Cooperative placement gains & Often assumes static helper relationships. \\
ICN and closeness caching \cite{Zhang2014NDN,Amadeo2022} & Connects content placement with network position & Usually not tied to WLAN handoff perturbation. \\
Mobility-aware caching \cite{Wang2017Mobility,Mardham2018} & Uses movement or contact patterns & Often relies on predictions or encounter heuristics rather than a direct AP handoff objective. \\
Recent vehicular/UAV/satellite caching \cite{bai2025collaborative,xu2026cooperative,ma2025delay,cao2025,zhao2026collaborative} & Models emerging mobile edge services & Usually targets vehicular, aerial, or satellite dynamics rather than AP-level WLAN handoff topology. \\
DRL caching \cite{mnih2015dqn,Min2020MARL,jiang2026cooperative} & Adapts to non-stationarity & Can be expensive if trained over the full placement space without a warm start. \\
Graph learning for networks \cite{Rusek2019,Jiang2022Survey,wang2024knowledge} & Models relational structure & Often requires heavier training or is not paired with a complete perturbation-aware caching pipeline. \\
\bottomrule
\end{tabular}
\end{table}
