\section{System Model and Problem Formulation}

\subsection{Network and Content Model}

We consider an IoT edge network with a set of edge caching nodes \(V\). In the WLAN setting, each node corresponds to an AP-like edge cache. The mobility-derived cooperation topology is represented by a weighted graph
\[
G=(V,E,W),
\]
where \(E\) is the set of observed cooperation relationships and \(W=\{w_{ij}\}\) is the interaction weight matrix. A larger \(w_{ij}\) indicates a stronger mobility-induced relationship between nodes \(i\) and \(j\), such as frequent handoffs between the corresponding APs. The graph is used both for cooperative retrieval latency and for redundancy modeling.

The system contains a content catalog \(\mathcal{C}\). Each node \(i\) has capacity \(K_i\), measured as the number of objects that can be stored. The binary placement variable is
\[
x_{i,c} =
\begin{cases}
1, & \text{if node } i \text{ stores content } c,\\
0, & \text{otherwise.}
\end{cases}
\]
The capacity constraint is
\[
\sum_{c \in \mathcal{C}} x_{i,c} \le K_i, \qquad \forall i \in V.
\]
Let \(P_{i,c}\) denote the probability that a request arriving at node \(i\) asks for content \(c\). The demand matrix \(P\) is node-specific, so two APs may have different content preferences even when the global catalog popularity is shared.

\subsection{Perturbed Cooperation Graph}

Real edge cooperation is uncertain. We model this uncertainty by sampling a perturbed graph
\[
\tilde{G}=(\tilde{V},\tilde{E},\tilde{W}) \sim \mathcal{P}_p(G),
\]
where \(p\) is the perturbation rate. In the experimental model, nodes are removed with probability \(p\), edges are removed with probability \(p/2\), and the largest connected component is retained for evaluation. Node removal represents temporary AP unavailability, maintenance, failure, or a time window in which an AP does not participate in cooperative caching. Edge removal represents a weakened cooperative relationship or loss of a usable path between APs.

Perturbation changes the value of a placement without changing the placement matrix itself. If a holder node is removed, its cached objects cannot serve active requests. If an edge is removed, shortest paths can become longer or disconnected. If a placement concentrates many useful objects at a few central nodes, it may perform well on the nominal graph but become fragile under node perturbation. If a placement spreads objects more broadly, it may trade nominal precision for perturbation robustness.

\subsection{Access Cost}

Let \(L_{\text{local}}\) be the latency of a local cache hit and \(L_{\text{origin}}\) be the latency of retrieving content from the origin. Let \(L(i,j;\tilde{G})\) be the cooperative retrieval latency from node \(j\) to node \(i\), computed as a non-decreasing function of shortest-path distance in the active perturbed graph. For a request for content \(c\) at node \(i\), the best non-local retrieval latency is
\[
A_{i,c}(x,\tilde{G})
=
\min\left(
L_{\text{origin}},
\min_{\substack{j \in \tilde{V}\\j \ne i\\x_{j,c}=1}} L(i,j;\tilde{G})
\right).
\]
The expected access cost is
\[
C_{\text{access}}(x,\tilde{G})
=
\sum_{i \in \tilde{V}}\sum_{c \in \mathcal{C}}
P_{i,c}
\left[
x_{i,c}L_{\text{local}}+
(1-x_{i,c})A_{i,c}(x,\tilde{G})
\right].
\]
This term rewards local hits and nearby cooperative hits while penalizing origin misses.

\subsection{Replication, Redundancy, and Relocation}

Replication cost captures storage and update overhead:
\[
C_{\text{rep}}(x,\tilde{G})
=
\sum_{i \in \tilde{V}}\sum_{c \in \mathcal{C}}
\gamma f_c x_{i,c},
\]
where \(f_c\) is a content-size or content-maintenance factor and \(\gamma\) is the unit replication weight. In the reported experiment, objects are treated uniformly, but the formulation allows heterogeneous objects.

Topology-aware redundancy penalizes storing the same object at strongly related active nodes:
\[
C_{\text{red}}(x,\tilde{G})
=
\sum_{c \in \mathcal{C}}\sum_{\substack{i,j \in \tilde{V}\\i<j}}
\tilde{w}_{ij}x_{i,c}x_{j,c}.
\]
Duplicate content is not prohibited. Highly popular objects may deserve multiple replicas. The redundancy term instead makes duplication pay a cost proportional to mobility-derived relatedness. Duplicating an object at APs that exchange many users is less useful than duplicating it at weakly related APs, because strong neighbors can often serve one another cooperatively.

Relocation cost measures how much a placement differs from a reference placement \(x^0\):
\[
C_{\text{rel}}(x,x^0)
=
\sum_{i \in \tilde{V}}\sum_{c \in \mathcal{C}}
\delta_c |x_{i,c}-x^0_{i,c}|.
\]
This term is important for adaptive caching. Without relocation cost, a learning component could appear to improve access cost by rewriting too much of the cache state. Relocation cost makes cache churn visible in the objective.

\subsection{Optimization Objective}

The stochastic topology-aware cooperative caching problem is
\[
\min_x
\mathbb{E}_{\tilde{G}\sim\mathcal{P}_p(G)}
\left[
C_{\text{access}}(x,\tilde{G})
+ \lambda C_{\text{rep}}(x,\tilde{G})
+ \beta C_{\text{red}}(x,\tilde{G})
+ \mu C_{\text{rel}}(x,x^0)
\right],
\]
subject to binary placement and cache-capacity constraints. The parameters \(\lambda\), \(\beta\), and \(\mu\) control the relative importance of replication, redundancy, and relocation. These weights are deployment-dependent: a network with cheap storage may tolerate more replicas, while a network with expensive updates may require stronger relocation control.

\subsection{Problem Properties}

The placement space is combinatorial. With \(N=|V|\) nodes, \(M=|\mathcal{C}|\) objects, and uniform cache capacity \(K\), the number of feasible placements grows exponentially with \(N\) and \(M\). Even without perturbation or relocation, cooperative placement resembles facility-location and coverage problems, which are known to be computationally challenging in wireless caching settings \cite{Poularakis2014Approx}. Perturbation adds another layer because the value of a placement is an expectation over graph samples rather than a deterministic value on one graph.

These properties motivate a hybrid method. Greedy placement provides a structured approximation by repeatedly selecting high marginal-gain node-content pairs. DQN refinement searches a local neighborhood around that greedy solution. Adaptive selection compares policy families under validation perturbations. The method therefore avoids both extremes: it does not rely only on a static heuristic, and it does not require a neural network to discover the entire combinatorial structure from scratch.

\subsection{Policy Families}

The formulation supports several policy families. Node-local policies use only \(P_{i,c}\) and ignore \(G\). Global popularity policies use aggregate demand but ignore topology. Diversity-aware policies increase catalog coverage by avoiding identical top-\(K\) placements everywhere. Topology-aware greedy policies optimize the stochastic objective directly. Learning-based policies refine placements based on sampled objective improvements. Adaptive TACC combines policy families by selecting the one that best matches the perturbation regime.
