\section{Methodology}

\subsection{Overview}

The proposed method is a three-stage online policy framework built on top of a mobility-derived graph, a node-specific demand matrix, and the currently deployed cache state. The first stage constructs strong candidate placements, including diversified popularity and topology-aware greedy placement. The second stage refines the currently deployed placement using DQN-based local search. The third stage is a lightweight validation gate that compares keep-current, diversified, greedy, and online DQN candidates using held-out perturbation samples, relocation cost against the deployed cache, variability, and computation cost. The method is designed around the observation that adaptation is useful only when the new placement improves the current cache enough to justify rewriting it.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    node distance=1.1cm,
    block/.style={draw, rounded corners, minimum width=3.5cm, minimum height=0.75cm, align=center},
    arrow/.style={-{Latex[length=2mm]}, thick}
]
\node[block] (trace) {Movement Trace};
\node[block, below=of trace] (graph) {Handoff Graph};
\node[block, below=of graph] (demand) {Zipf-Locality Demand};
\node[block, below left=0.7cm and 1.7cm of demand] (greedy) {Topology-Aware Greedy};
\node[block, below=0.7cm of demand] (diverse) {Diversified Placement};
\node[block, below right=0.7cm and 1.7cm of demand] (dqn) {Online DQN Refinement};
\node[block, below=2.0cm of demand] (selector) {Validation Gate};
\node[block, below=of selector] (eval) {Perturbation Evaluation};
\draw[arrow] (trace) -- (graph);
\draw[arrow] (graph) -- (demand);
\draw[arrow] (demand) -- (greedy);
\draw[arrow] (demand) -- (diverse);
\draw[arrow] (demand) -- (dqn);
\draw[arrow] (greedy) -- (selector);
\draw[arrow] (diverse) -- (selector);
\draw[arrow] (dqn) -- (selector);
\draw[arrow] (selector) -- (eval);
\end{tikzpicture}
\caption{Online topology-aware caching pipeline. Candidate placements are evaluated under graph perturbation against the currently deployed cache state, and the validation gate selects a deployment using cache objective, relocation cost, variability, and optimization cost.}
\label{fig:pipeline}
\end{figure}

\subsection{Graph Construction}

The mobility trace is represented as sequences of AP associations. Consecutive AP changes by the same client create handoff events. Handoff counts are aggregated over all clients to form a weighted undirected graph. Each edge stores a raw handoff count, a normalized redundancy weight, and an inverse-count latency weight. Frequent handoffs therefore imply strong relatedness and low cooperative retrieval cost. The largest active component is used for evaluation so that cooperative retrieval is defined over a connected graph.

This graph construction makes the cooperation topology data-dependent. Two APs are adjacent because clients move between them, not because a fixed radius or manually chosen neighborhood says they should cooperate. The resulting topology captures the behavioral structure that is relevant for caching: where demand may migrate and where cooperative retrieval is likely to be useful.

\subsection{Demand Construction}

Content demand is generated using a global Zipf distribution and AP-level locality. Let \(q_c\propto c^{-\alpha}\) denote global catalog popularity. For each AP community, a subset of objects receives a locality boost, after which probabilities are normalized to produce \(P_{i,c}\). This creates demand distributions that share a global popularity shape but differ across APs. The locality multiplier controls how strongly node demand deviates from the global distribution.

This demand model is not intended to replace real content logs. It is a controlled way to evaluate caching behavior when the mobility trace contains association events but not object requests. The model also supports a meaningful comparison among local popularity, global popularity, diversified placement, and topology-aware placement because the policies face both common popular objects and location-specific preferences.

\subsection{Topology Perturbation}

For perturbation rate \(p\), each node is removed with probability \(p\), and each edge is removed with probability \(p/2\). The largest connected component is retained. This sampler creates progressively less stable cooperative graphs while preserving enough structure for evaluation. Candidate policies are evaluated over repeated samples so that objective mean and variance can be measured.

The perturbation model is intentionally simple, but it isolates an important effect. A placement optimized for the nominal graph may rely on central nodes or short paths that disappear under perturbation. A more diverse placement may survive moderate perturbation better because more objects remain reachable. The selector uses this difference to choose among candidate policies and to decide whether DQN refinement is worth its added cost.

The perturbation rate should be interpreted as an operating-regime descriptor rather than as a permanent physical failure probability. A larger \(p\) means that the controller sees fewer active APs and fewer reliable cooperative links in a validation window. This changes access paths, origin misses, redundancy relationships, and the value of replicas. The experiments evaluate \(p\) through 0.25 to test whether stronger perturbation makes the validation gate useful or whether transition-aware online DQN remains sufficient. Beyond this range, many sampled graphs can become dominated by disconnected or near-local behavior, and the system should trigger topology re-estimation or a fallback placement mode rather than assuming that the same cooperative objective remains equally informative.

\subsection{Baseline Policies}

The baseline set is designed to avoid weak comparisons. Random placement measures the value of unstructured diversity. Local popularity measures independent AP optimization. Global popularity measures the effect of storing the same popular objects everywhere. Diversified popularity shifts popular-object slices across APs to increase cooperative catalog coverage. Topology-aware greedy measures direct optimization of the proposed objective. Fixed Greedy+DQN measures learning-based refinement around a structured warm start. Online DQN measures transition-aware refinement from the currently deployed cache. The validation-gated row measures whether a cost-aware safeguard changes the deployment relative to Online DQN.

\subsection{Topology-Aware Greedy Placement}

The greedy algorithm starts with an empty placement and repeatedly adds the feasible node-content pair with the largest expected objective reduction. Candidate gains are estimated over sampled perturbed graphs:
\[
\Delta(i,c|x)=
\mathbb{E}_{\tilde{G}\sim\mathcal{P}_p(G)}
\left[J(x,\tilde{G})-J(x^{(i,c)},\tilde{G})\right],
\]
where \(x^{(i,c)}\) is the placement obtained by adding content \(c\) to node \(i\). Candidate contents are restricted to high-demand objects at each node to keep marginal-gain evaluation tractable.

This stage is one of the main contributions because the marginal gain is not computed from popularity alone. A candidate object is valuable when it reduces local or cooperative access cost, avoids unnecessary origin retrieval, and does not create excessive duplicate storage among strongly related APs. The same object may therefore be attractive at a weakly connected AP but unattractive at a strongly related neighbor that can already retrieve it cheaply. Sampling perturbed graphs inside the gain estimate also discourages placements that depend too heavily on a single central holder.

The candidate restriction is a computational device, not a change in the optimization objective. For each AP, the algorithm considers objects with high local demand and high global demand, then scores those feasible pairs using the full objective. This preserves the important object choices while avoiding a full scan of the \(N|\mathcal{C}|\) action space at every greedy step. Since each node has finite cache capacity, the loop stops after no feasible pair provides positive expected reduction or after all cache slots are filled.

\begin{algorithm}[t]
\caption{Topology-Aware Greedy Placement}
\label{alg:greedy_final}
\begin{algorithmic}[1]
\Require Graph \(G\), demand \(P\), capacity \(K\), perturbation sampler \(\mathcal{P}_p\)
\State Initialize placement \(x\gets 0\)
\While{some node has free cache capacity}
    \State bestGain \(\gets 0\), bestPair \(\gets \varnothing\)
    \For{each feasible node-content pair \((i,c)\)}
        \State Estimate \(\Delta(i,c|x)\) over sampled perturbations
        \If{\(\Delta(i,c|x)>\) bestGain}
            \State bestGain \(\gets \Delta(i,c|x)\), bestPair \(\gets (i,c)\)
        \EndIf
    \EndFor
    \If{bestPair is empty}
        \State \textbf{break}
    \EndIf
    \State Add bestPair to placement \(x\)
\EndWhile
\State \Return \(x\)
\end{algorithmic}
\end{algorithm}

\subsection{DQN Refinement}

The DQN refiner treats cache modification as a local sequential decision problem. Unlike a fixed DQN placement that always starts from topology-aware greedy, the online refiner starts from the currently deployed cache state \(x^{t-1}\). The state concatenates the current binary placement matrix, the demand matrix, and graph features. Graph features include normalized degree, weighted degree, and closeness centrality. The action space is a node-content toggle. If adding an object to a full cache would violate capacity, the lowest-demand cached object at that node is evicted. The reward is online objective improvement:
\[
r_t=J(x_t,x^{t-1},\tilde{G}_t)-J(x_{t+1},x^{t-1},\tilde{G}_t).
\]

The DQN uses two hidden layers with ReLU activations, replay-buffer sampling, a target network, discount factor 0.92, learning rate 0.001, and epsilon-greedy exploration. During inference, an action is accepted only if it improves the measured online objective averaged over perturbation samples. This accepted-action rule is critical because it prevents the learned component from degrading the deployed cache state. In effect, the neural network proposes local modifications, and the objective acts as a safety filter.

The learning component is therefore deliberately bounded. It is not trained to construct a full placement from scratch, which would make the action space too large and the result difficult to interpret. Instead, the DQN learns where the currently deployed solution has local weaknesses under sampled perturbations. A useful action may replace a redundant object, introduce a replica at a more resilient node, or reduce expected cooperative path length. If the proposed move fails the objective check, the placement remains unchanged. This design makes learning part of the adaptive method while preserving relocation awareness.

\subsection{Validation Gate}

The validation gate selects among keep-current, diversified popularity, topology-aware greedy, and online DQN refinement. It evaluates each candidate against the currently deployed placement \(x^{t-1}\), so a candidate that requires many cache rewrites must earn enough access-cost or redundancy improvement to justify its relocation cost. Online DQN is trained from \(x^{t-1}\), which gives learning a concrete role: it searches for low-churn local improvements rather than proposing a fixed replacement placement.

The gate answers a different question from the placement algorithms. Keep-current asks whether rewriting can be avoided. Diversified popularity asks whether broad catalog coverage is enough. Topology-aware greedy asks whether directly optimizing the stochastic objective is enough. Online DQN asks whether local learned substitutions improve the deployed cache with acceptable churn. The gate asks which of those answers is justified in the current perturbation regime after accounting for validation quality, variability, replication pressure, relocation, and optimization overhead.

For each candidate policy, the selector computes
\[
S(\pi)=\bar{J}_{\text{val}}(\pi)+\alpha\sigma_{\text{val}}(\pi)+\rho p^2 R(\pi)+\eta\Omega(\pi)-\kappa U(\pi),
\]
where \(\bar{J}_{\text{val}}\) is validation objective, \(\sigma_{\text{val}}\) is validation standard deviation, \(R(\pi)\) is average replication cost, \(\Omega(\pi)\) is a normalized optimization-cost estimate, and \(U(\pi)\) is an instability credit applied to DQN when the best cheap candidate is highly variable. The \(p^2\) term makes replication pressure grow smoothly with perturbation intensity rather than using a fixed regime switch. The selected policy is
\[
\pi^*=\arg\min_{\pi}S(\pi).
\]

This rule keeps selection dynamic rather than fixed. If a cheap placement is already reliable, it can be selected without paying learning cost. If the current cache has local weaknesses, online DQN can be selected because it modifies the deployed placement incrementally rather than replacing it wholesale. The policy choice is therefore learned from validation evidence in each regime, not hard-coded by perturbation rate.

\begin{algorithm}[t]
\caption{Validation-Gated Policy Selection}
\label{alg:adaptive_selector}
\begin{algorithmic}[1]
\Require Cheap policies \(\Pi_0\), deployed placement \(x^{t-1}\), perturbation rate \(p\), validation sampler \(\mathcal{V}_p\)
\For{each policy \(\pi \in \Pi_0\)}
    \State Evaluate \(\pi\) against \(x^{t-1}\) on held-out perturbation samples from \(\mathcal{V}_p\)
    \State Compute cost-aware score \(S(\pi)\)
\EndFor
\State Train online DQN from \(x^{t-1}\) and produce refined placement \(\pi_D(x^{t-1})\)
\State Compute \(S(\pi_D)\) including relocation, optimization cost, and instability credit
\State \Return policy with minimum computed score \(S(\pi)\)
\end{algorithmic}
\end{algorithm}

\subsection{Computational Complexity}

The greedy stage is the most expensive deterministic component. If there are \(N\) nodes, \(M\) content objects, cache capacity \(K\), and \(S\) perturbation samples per marginal evaluation, naive greedy search can require \(O(NKMNS)\) objective evaluations. Candidate filtering reduces the effective value of \(M\) by considering only high-demand objects at each node. The DQN action space has size \(NM\), and training cost scales with the number of episodes, steps per episode, and minibatch size. Since online DQN starts from the deployed placement, it does not need to learn an entire placement from an empty cache state.

The gate adds validation overhead for a small set of candidate policies. In the final run, online DQN is selected in all six perturbation windows, including \(p=0.20\) and \(p=0.25\). This result is important because it prevents overclaiming: the gate is not the source of the measured gains in this experiment. It gives DQN a specific operational role as the low-churn online improvement operator, while keep-current, diversified, and greedy placements remain strong non-neural candidates and sanity checks.
