\section{Discussion}

The main finding is that topology awareness is valuable but not sufficient by itself. The topology-aware greedy policy substantially improves over local and global popularity because it uses the handoff graph to place content where cooperative retrieval is useful. Diversified popularity is also a strong non-neural candidate because it increases catalog coverage and remains competitive under perturbation. However, the sequential online setting changes the main question: the controller must decide whether changing the currently deployed cache is worth the relocation cost. In that setting, Online DQN is the strongest learned refinement policy because it refines the deployed placement incrementally rather than forcing a wholesale switch to a separate static policy. The \(p=0.25\) result shows that this refinement should still be compared against a no-update option before rewriting the cache.

The DQN result should be interpreted as transition-aware local refinement, not as end-to-end neural optimization. The full placement space is too large for a small DQN to learn from scratch reliably. Starting from the deployed placement changes the problem: the neural agent only needs to identify beneficial local substitutions. This is why Online DQN is stronger than fixed Greedy+DQN at every tested perturbation rate. Fixed Greedy+DQN improves a structured warm start, but Online DQN optimizes the actual transition the controller faces; the \(p=0.25\) Keep Current result shows that this transition should still be accepted only when it beats no-update.

The relocation term plays an important stabilizing role. Without relocation cost, a learning agent could reduce access cost by frequently rewriting caches. That behavior may be unrealistic in IoT edge systems where update bandwidth, storage writes, and control overhead are limited. Online DQN reports nonzero relocation precisely because it is evaluated as a sequential controller. The measured relocation grows from 0.150 objects per active node in the stable window to 1.113 at \(p=0.20\) and 1.022 at \(p=0.25\), while objective remains much lower than fixed topology-greedy and fixed Greedy+DQN alternatives at high perturbation. This supports the interpretation that targeted online modification is valuable when its cost is explicitly charged.

This interpretation also answers what should happen when the input changes. If demand shifts, the demand matrix changes the marginal gains and DQN rewards. If mobility changes, the handoff graph changes shortest paths, edge weights, redundancy penalties, and perturbation samples. If the perturbation rate increases, the online refiner sees a different objective landscape and different relocation trade-offs. The controller is therefore not a fixed rule that maps one \(p\) value to one placement forever; it is an online refinement procedure that can be rerun when the observed graph, demand, or stability window changes.

The results also show why strong baselines are necessary. If the evaluation only compared against random, local popularity, and global popularity, topology-aware greedy would appear uniformly dominant. Diversified popularity reveals a subtler conclusion: a simple diversity mechanism can outperform graph-aware greedy under some perturbation settings when viewed as a continuously deployed fixed policy. %This does not weaken the contribution; it clarifies it. The contribution is the online combination of structural placement and transition-aware refinement.This result refines the interpretation of the contribution. 
The main contribution is not topology-aware greedy placement alone, but the online combination of structural placement and transition-aware refinement.

The objective weights also deserve deployment-specific interpretation. If storage is cheap and content updates are infrequent, a system may prefer higher diversity and tolerate more replication. If update bandwidth is scarce or caches have limited write endurance, relocation and replication penalties should be stronger. If origin latency is extremely high, the access-cost term should dominate. The proposed objective is therefore a modeling framework whose weights should be calibrated for the deployment rather than treated as universal constants.

The use of trace-compatible mobility data limits claims about real-world magnitude, but it does not invalidate the methodological conclusion. This study evaluates how policy families behave under a mobility-derived graph, heterogeneous demand, perturbation, and sequential cache updates. The qualitative pattern is coherent: exact graph-aware placement provides a strong initial state, diversity provides a strong non-neural comparison, and transition-aware DQN provides the strongest learned refinement while the no-update baseline remains necessary under severe perturbation.

The \(p=0.20\) and \(p=0.25\) cases should be treated as stronger stress conditions than the low-perturbation windows. They are included in the final results because they test whether online refinement remains useful as perturbation increases. The no-update row is competitive at \(p=0.25\), which is a useful caution: the paper should not claim that cache rewriting is always beneficial. Larger perturbation beyond \(p=0.25\) may leave too little connected structure for cooperative caching to dominate local service. In such a regime, the practical response is to detect loss of cooperative structure, reduce reliance on remote holders, and re-optimize after enough new mobility observations arrive.

The cost terms should also be interpreted as guardrails rather than as the dominant source of objective variation. Access cost is deliberately dominant because the main system goal is to reduce retrieval delay and origin reliance. Replication, redundancy, and relocation are weighted softly so that the controller can still rewrite caches when the access benefit is clear. If a deployment has expensive cache updates or limited write endurance, \(\mu\) should be increased and Online DQN should require a stronger improvement margin before accepting cache edits.
