\section{Limitations and Threats to Validity}

The first limitation is dataset access. The study is designed around the Dartmouth campus WLAN trace, but the raw movement archive was not directly available through legacy unauthenticated URLs at the time of this work. The evaluation therefore uses deterministic trace-compatible mobility data. This validates the pipeline but does not replace evaluation on the raw Dartmouth archive.

The second limitation is synthetic content demand. Zipf popularity and spatial locality are research-backed assumptions, but real application demand may include burstiness, temporal drift, content lifetimes, and user-specific preferences. Future work should rerun the pipeline with real content logs if available.

The third limitation is scale. The final experiment uses 20 nodes and 48 objects. Larger graphs would require faster candidate evaluation, batched shortest-path computation, and possibly graph neural encoders. The current method favors clarity and reproducibility over large-scale optimization.

The fourth limitation is online acceptance under high perturbation. The \(p=0.25\) window illustrates this risk because Keep Current is slightly better on observed mean objective than Online DQN, even though Online DQN remains much stronger than fixed Greedy+DQN and topology-aware greedy. This should be addressed with online monitoring, confidence intervals, and stricter accepted-action thresholds.

The fifth limitation is perturbation range. The experiments evaluate \(p\leq 0.25\), where the graph remains degraded but cooperative caching is still meaningful in the sampled runs. Higher perturbation rates should be studied as a separate failure-mode setting with topology re-estimation, fallback placement, and recovery policies.

Finally, the DQN uses explicit graph features rather than a full GNN. This keeps the model lightweight, but a GNN may generalize better across graph sizes and node permutations. The present method should therefore be seen as a strong baseline and research scaffold for future graph-learning extensions.
