\section{Future Work}

Several extensions would strengthen the present study:

\begin{itemize}
    \item \textbf{Raw-trace evaluation:} Evaluate the same pipeline on the raw Dartmouth/IEEE DataPort WLAN archive or another public AP-association trace. The current trace-compatible generator preserves the pipeline and creates meaningful handoff structure, but raw mobility traces would allow stronger claims about deployment magnitude, temporal stationarity, and AP heterogeneity.
    \item \textbf{Temporal online evaluation:} Split mobility into time windows, estimate the handoff graph and demand model from past windows, deploy a cache placement, and evaluate on future windows. This would test whether Online DQN remains useful when graph drift, demand drift, and perturbation are observed rather than sampled.
    \item \textbf{Online acceptance calibration:} Add confidence intervals, sequential statistical tests, and explicit no-update thresholds. The final sweep shows that Keep Current is slightly better on observed mean objective at \(p=0.25\); an online controller should avoid rewriting when the expected gain is not statistically reliable.
    \item \textbf{Cost-weight sensitivity:} Sweep \(\lambda\), \(\beta\), and \(\mu\) to quantify when replication, redundancy, and relocation become decisive. This is especially important for deployments with expensive cache writes, limited storage endurance, or strict backhaul-update budgets.
    \item \textbf{Larger-scale experiments:} Scale beyond 20 APs and 48 content objects using batched shortest-path computation, more efficient marginal-gain screening, and possibly distributed training. The online DQN action space grows with \(N|\mathcal{C}|\), so larger deployments may need hierarchical actions, candidate pruning, or graph-based action embeddings.
    \item \textbf{Graph neural policies:} Replace explicit degree, weighted-degree, and closeness features with a permutation-aware GNN encoder that generalizes across topologies. Such a model should still be evaluated against diversified placement, topology-aware greedy, Keep Current, and transition-aware Online DQN, and it should continue to charge relocation cost against the previous deployed cache.
\end{itemize}
